---
title: "Describe Data - Pier"
output: html_notebook
---

> Do we need to segregate the matches by type?   
> How does type influence the match gameplay? 

golddiff is defined as goldblue - goldred, by the minute  

> Can we can remove goldblue, goldred? Is the actual quantity significant? 

```{r}
match_categories <- unique(matches$Type)
print(match_categories)
print(length(match_categories))



```
```{r}
# extract just golddiff for analysis
just_gold_diff = gsub("\\[|\\]", "", matches$golddiff)
golddiff_split_by_min = do.call("rbind", strsplit(just_gold_diff, ",")) # remove the , 
golddiff_split_by_min = data.frame(apply(golddiff_split_by_min, 2, as.numeric)) # convert str to num
head(golddiff_split_by_min)
```
```{r}
plot(colMeans(golddiff_split_by_min), type="b", xlab="Minutes", ylab="Average Gold Diff (Blue - Red)")
```

> Insanely skewed towards blue! No negative values (none indicating red had more gold than blue) at all. Did I miss something? 

The following is an exercise using *only* TOP position data. 

```{r}
unique(matches$blueTopChamp)
with(matches, cor(as.numeric(matches$blueTop), as.numeric(matches$blueTopChamp)))
```
There are 75 different champions used in the *top* position. 
There is not a strong correlation between the player and the champion that they use in the matches. 
```{r}
library(ggplot2)
t <- table(matches$blueTopChamp) # gets frequencies
t_sorted = sort(t, decreasing=TRUE)[1:10] # sorts and gets top 10
t_sorted_df = data.frame(t_sorted)
names(t_sorted_df)[names(t_sorted_df) == 'Var1'] <- 'Champion'
ggplot(t_sorted_df, aes(x=Champion,y =Freq)) +geom_bar(stat = "identity") + ggtitle("Top 10 Champions Used (Top Position)") + theme(plot.title = element_text(hjust = 0.5))

```
Now the bottom 10 Champions. (least frequently used)
```{r}
t_sorted = sort(t, decreasing=FALSE)[1:11] # sorts and gets bottom 11
t_sorted_df = data.frame(t_sorted)
names(t_sorted_df)[names(t_sorted_df) == 'Var1'] <- 'Champion'
ggplot(t_sorted_df, aes(x=Champion,y =Freq)) +geom_bar(stat = "identity") + ggtitle("Bottom 10 Champions Used (Top Position)") + theme(plot.title = element_text(hjust = 0.5))
```



