---
title: "League of Legends"
output:
  html_document:
    keep_md: true
---
Research question: see dreamteam files
========================================================
#### Table of content (type of analysis)
Section 0: preparing data (no results shown)  
Section 1: exploring variables and relations (mostly univariate and some
bivariate)  
Section 2: exploring differnces between results (won/vs lost) and control (bi- and multivariate)  
Section 3: other relations between variables (bi- and multivariate)  
Section 4: Further exploratory analysis to understand game dynamics

> **Data source**: League of Legends competitive matches between 2015-2017. The matches include the NALCS, EULCS, LCK, LMS, and CBLoL leagues as well as the World Championship and Mid-Season Invitational tournaments. https://www.kaggle.com/chuckephron/leagueoflegends

```{r "0.1 packages"}
library(ggplot2)
library(tidyr)
library(stringr)
library(dplyr)

```



```{r "0.3 Functions"}

# 0.3.1 cuberoot
cuberoot_trans = function() trans_new('cuberoot', 
                                      transform = function(x) x^(1/3), 
                                      inverse = function(x) X^3)

# 0.3.2 Define new column names: takes in the max number for column and a name, outputs multiple columns
# Problem: appending to an object in a for loop causes the entire object to be copied on every iteration --> very slow.
create_column_names <- function(columns, namestring){
  maxnr <- max(matches$columns)
  vectofnr = c();
  for (i in 1:81){
    # print(i)
    vectofnr[i] <- paste(namestring, as.character(i))
    # print(vectofnr)
  }
  return(vectofnr)
}
# because it is so slow, maybe you can also put it in a global var:
# listofcolumns <- create_column_names(gamelength)

# 0.3.3 remove the first and last char
remove_first_last <- function(inputstring){
  return(substring(inputstring, 2, str_length(inputstring)-1))
}

```


```{r "0.4 Data Wrangling"}

# 0.4.1 viewing & correcting types
str(matches)
# Types seem ok, but have to correct the lists.

## 0.4.2 making interpretation (and graphs) more easy by adding winner (blue or red) column
matches$winner <- ifelse(matches$bResult > 0, c("blue"), c("red"))

## 0.4.3 get rid of [ & ] for later use --> only use once!
matches$goldblue <- remove_first_last(matches$goldblue)
matches$golddiff <- remove_first_last(matches$golddiff)
matches$goldred <- remove_first_last(matches$goldred)


## 0.4.4 Square the data if scewed 

# to do

```


```{r "0.5 New variables"}
matches$winner <- ifelse(matches$bResult > 0, c("blue"), c("red"))

# matches$YearSeason <- as.character(matches$Year)
matches$YearSeason <- ifelse(matches$Season == "Spring", paste("01/01/", as.character(matches$Year), sep=""), paste("01/06/", as.character(matches$Year), sep=""))

str(matches$YearSeason)
summary(matches$Season)


matches$YearSeason <- as.Date(matches$YearSeason, format = "%d/%m/%Y")

summary(matches$YearSeason)
```

```{r "0.6 Making subsets"}
# It gets messy very easily so extra dataset instead of extra variable:
matches_seperated <- matches %>%
  separate(golddiff, into = c(create_column_names(gamelength, "gold diff in minute")), sep = "\\,")

# may be interesting to make a red vs blue won dataset.

blueset <- subset(matches, winner == 'blue')
redset <- subset(matches, winner == 'red')
```

# 1. EXPLORING THE DATA SET (univariate and some bivariate).

## 1.1 First exploratory plots (see below graph for analysis)

```{r}
# Get the names
names(matches)
# 58 original vars

str(matches)
# So there are 185 teams --> seems like a lot? Do teams change from name? Do members change?

# this runs a whole summary (lot of work):
summary(matches)

table(matches$winner)
# ---> so blue wins (a lot) more
```

Observations:
Regional competition is very small: 143... exclude those if they behave very differently

There are two seasons, so we seem to have Year*Season = 4*2=8 time values

Game length mean and median are very close: so it is not scewed.

Most notable is the choice for Maokai for top red--> he is a lot more favorite for red than blue...

```{r results = 'hide'}
# make a table of RG Case --> it is exactly the same, so that seems correct
table(matches$winner)
#it is exactly the same, so that seems correct
```
## 1.2 First exploratory plots: length of the game and the different leagues, types & timeframes(seasons)

How are the leagues distributed (univariate)?  
```{r "1.2 First exploratory plots"}
qplot(x = League, data = matches,
      xlab = 'league',
      ylab = 'Number of games from sample',
      color = I('black'), fill = I('blue'))
```
A lot of variation, with clear distinction. The game should function the same across the leagues.

How are the types distributed (univariate)?  
```{r}
qplot(x = Type, data = matches,
      xlab = 'league',
      ylab = 'Number of games from sample',
      color = I('black'), fill = I('yellow'))
summary(matches$Type)
```
Huge difference between season and regional --> see if regional behaves differently, maybe exclude from sample

How are the seasons distributed (univariate)?  
```{r}
qplot(x = Type, data = matches,
      xlab = 'league',
      ylab = 'Number of games from sample',
      color = I('black'), fill = I('yellow'))
summary(matches$Type)
```





How is the time distributed (univariate)?  
```{r}
qplot(x = gamelength, data = matches,
      xlab = 'length of game',
      ylab = 'Number of games from sample',
      color = I('black'), fill = I('purple')) +
  scale_x_continuous()
```
Like we saw from the summary, the data is nicely distributed. It is also clear that it is almost impossible to win this game in the first 20 minutes in the professional spfere.

How are the matches distributed of the seasons (univariate)?  
```{r}
qplot(x = YearSeason, data = matches,
      xlab = 'Season',
      ylab = 'Number of games from sample',
      color = I('black'), fill = I('purple'))
```
Wow, there is way more data from later timeslots...

How is the time distributed over the different leagues (bivar)?

```{r}
ggplot(aes(x = gamelength, y = YearSeason), data = matches) + 
  geom_point(alpha = 1/6, color = 'red', position = position_jitter(h = 0)) +
  xlim(16, 80) 
```
Not much to see, the games are getting longer, but also much more data so no real conclusion to be made
```{r}
theme_set(theme_minimal(20))
set.seed(1836)
names(matches_seperated)
golddiff_subset <- matches_seperated[,c("gold diff in minute 1",  "gold diff in minute 2",  "gold diff in minute 3", 
 "gold diff in minute 4",  "gold diff in minute 5",  "gold diff in minute 6",  "gold diff in minute 7", 
 "gold diff in minute 8",  "gold diff in minute 9",  "gold diff in minute 10", "gold diff in minute 11",
 "gold diff in minute 12", "gold diff in minute 13", "gold diff in minute 14", "gold diff in minute 15",
 "gold diff in minute 16", "gold diff in minute 17", "gold diff in minute 18", "gold diff in minute 19",
 "gold diff in minute 20")]
# summary(gam_subset)
# ggpairs(gam_subset[sample.int(nrow(gam_subset), 1000),]) +
#   theme(
#     axis.text = element_text(size = 3),
#     axis.title = element_text(size = 2),
#     legend.background = element_rect(fill = "white"),
#     panel.grid.major = element_line(colour = NA),
#     panel.grid.minor = element_blank(),
#     panel.background = element_rect(fill = "grey95"))

# gam_subset2 <- gam[,c("RG_case", 
#                       "age_at_registration",
#                       "frequency_fixedodds",
#                       "bets_per_day_fixedodds",
#                       "euros_per_bet_fixedodds",
#                       "net_loss_fixedodds",
#                       "percent_lost_fixedodds",
#                       "all_bets",
#                       "all_stakes",
#                       "Registration_Date_Year.bucket",
#                       "Registration_Date_Year.largebucket")]
# # ggpairs(gam_subset[sample.int(nrow(gam_subset), 1000),]) +
# #   theme(
# #     axis.text = element_text(size = 3),
# #     axis.title = element_text(size = 2),
# #     legend.background = element_rect(fill = "white"),
# #     panel.grid.major = element_line(colour = NA),
# #     panel.grid.minor = element_blank(),
# #     panel.background = element_rect(fill = "grey95"))
# 
# # only numerical:
# 
# gam_subset3 <- gam[,c("age_at_registration",
#                       "frequency_fixedodds",
#                       "bets_per_day_fixedodds",
#                       "euros_per_bet_fixedodds",
#                       "net_loss_fixedodds",
#                       "percent_lost_fixedodds",
#                       "all_bets",
#                       "all_stakes",
#                       "Registration_Date_Year.bucket",
#                       "Registration_Date_Year.largebucket",
#                       "Missing_Daily_Transactions",
#                       "First_Deposit_Date",
#                       "Registration_date",
#                       "sum_stakes_fixedodds", 
#                       "bettingdays_fixedodds",
#                       "duration_fixedodds")]


## Multiple t-tests for time, age, frail between genders.
lapply(gam[,c("bettingdays_fixedodds", "sum_stakes_fixedodds", "bets_per_day_fixedodds")], function(x) t.test(x ~ gam$RG_case))

ggcorr(golddiff_subset[sample.int(nrow(golddiff_subset), 1000),])
```
## 4  Further exploratory analysis to understand game dynamics

### 4.1 Does having more gold than your opponent for a longer period have an influence on the game outcome?
```{r}

matches_data = read.csv("data/matches.csv")


# Get the gold difference as a list of numbers - process the values by removing the [, ] brackets
getGoldDiffAsListOfNumbers <- function(matchNumber){
  # get the characters from the first row of the matches data and the golddiff column
  golddiffStr <- as.character(matches_data[matchNumber, c("golddiff")])
  
  # split the value by the comma
  golddiffList <- strsplit(golddiffStr,", ")
  
  # get the length of the number of entries in the gold difference list
  playLength <- length(golddiffList[[1]])
  
  # get the numbers equivalent to the elements
  golddiffListValues <- golddiffList[[1]]
  
  # replace [ with empty
  golddiffListValues[1] <- gsub("\\[", "", golddiffListValues[1])
  
  # replace ] with empty
  golddiffListValues[playLength] <- gsub("]", "", golddiffListValues[playLength])
  golddiffListValues[playLength] 
  
  # Convert values to numeric
  golddiffNumericList <- as.numeric(as.character(golddiffListValues))
  
  # return the list obtained
  golddiffNumericList
}

for (matchNumber in 1:nrow(matches_data)) {
  golddiffNumericList <- getGoldDiffAsListOfNumbers(matchNumber)
  timesBlueWithMoreGoldThanRed <- length(which(golddiffNumericList > 0))
  matchLength <- length(golddiffNumericList)
  
  # fraction of the game where blue team was leading in gold
  matches_data$ratioBlueMoreGoldThanRed[matchNumber] <- timesBlueWithMoreGoldThanRed/matchLength
}

timesBlueTeamWonForRatio = list()
timesBlueTeamLoseForRatio = list()

for (i in 1:10){
  ratioOfMatchBlueLedInGold <- 0.1 * i
  timesBlueTeamWonForRatio[i] <- length(which(matches_data$bResult[matches_data$ratioBlueMoreGoldThanRed > (ratioOfMatchBlueLedInGold - 0.1) & matches_data$ratioBlueMoreGoldThanRed <= ratioOfMatchBlueLedInGold] == 1))
  timesBlueTeamLoseForRatio[i] <- length(which(matches_data$bResult[matches_data$ratioBlueMoreGoldThanRed > (ratioOfMatchBlueLedInGold - 0.1) & matches_data$ratioBlueMoreGoldThanRed <= ratioOfMatchBlueLedInGold] == 0))
}

# scatter plot of the 2 different values
xValues <- seq(10, 100, by = 10)
plot(xValues, timesBlueTeamWonForRatio, type="p", main="Advantage in gold possession with winning", xlab="Percent of total match Blue team had more gold than Red team", ylab="Number of matches won", col="blue")
par(new = TRUE)
plot(xValues, timesBlueTeamLoseForRatio, type="p", col="red", axes = FALSE, xlab="", ylab="")
axis(1, at = xValues)
legend("top", legend = c("Matches won by Blue team", "Matches won by Red team"), col = c("blue", "red"), cex = 0.8, pch=1)



```