---
title: "League of Legends"
output:
  html_document: default
  pdf_document: default
---
Research question: see dreamteam files
========================================================
#### Table of content (type of analysis)
Section 0: preparing data (no results shown)  
Section 1: exploring variables and relations (mostly univariate and some
bivariate)
Section 2: exploring differnces between results (won/vs lost) and control (bi- and multivariate)  
Section 3: other relations between variables (bi- and multivariate)  


> **Data source**: League of Legends competitive matches between 2015-2017. The matches include the NALCS, EULCS, LCK, LMS, and CBLoL leagues as well as the World Championship and Mid-Season Invitational tournaments. https://www.kaggle.com/chuckephron/leagueoflegends

```{r echo=FALSE, message=FALSE, warning=FALSE, "0.1 packages"}

library(ggplot2)
library(bitops)
library(tidyr)
library(RCurl)
library(ggthemes)
library(foreign)
library(grid)
library(RColorBrewer)
library(gridExtra)
library(reshape2)
library(stringr)
library(scales)
library(dplyr)
library(GGally)
library(memisc)
library(lattice)
library(MASS)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, "0.2 Load_the_Data"}
# Source: https://www.kaggle.com/chuckephron/leagueoflegends
getwd()

# Tip: don't run seperately with load command.
# PC:
# setwd('C:/Users/Gebruiker 1/Dropbox/NUS/Data mining/leagueoflegends')
# laptop:
setwd('C:/Users/User/Dropbox/NUS/Data mining/leagueoflegends')

# load data file
matches <- read.csv("matches.csv", header=TRUE, sep=",")
```


```{r echo=FALSE, message=FALSE, warning=FALSE, "0.3 Functions"}

# 0.3.1 cuberoot
cuberoot_trans = function() trans_new('cuberoot', 
                                      transform = function(x) x^(1/3), 
                                      inverse = function(x) X^3)

# 0.3.2 Define new column names: takes in the max number for column and a name, outputs multiple columns
# Problem: appending to an object in a for loop causes the entire object to be copied on every iteration --> very slow.
create_column_names <- function(columns, namestring){
  maxnr <- max(matches$columns)
  vectofnr = c();
  for (i in 1:81){
    # print(i)
    vectofnr[i] <- paste(namestring, as.character(i))
    # print(vectofnr)
  }
  return(vectofnr)
}
# because it is so slow, maybe you can also put it in a global var:
listofcolumns <- create_column_names(gamelength)

# 0.3.3 remove the first and last char
remove_first_last <- function(inputstring){
  return(substring(inputstring, 2, str_length(inputstring)-1))
}

```


```{r echo=FALSE, message=FALSE, warning=FALSE, results = "hide", "0.4 Data Wrangling"}

# 0.4.1 viewing & correcting types
str(matches)
# Types seem ok, but have to correct the lists.

## 0.4.2 making interpretation (and graphs) more easy by adding winner (blue or red) column
matches$winner <- ifelse(matches$bResult > 0, c("blue"), c("red"))

## 0.4.3 get rid of [ & ] for later use --> only use once!
matches$goldblue <- remove_first_last(matches$goldblue)
matches$golddiff <- remove_first_last(matches$golddiff)
matches$goldred <- remove_first_last(matches$goldred)


## 0.4.4 Square the data if scewed 

# to do

```


```{r echo=FALSE, warning=FALSE, results = "hide", "0.5 New variables"}
matches$winner <- ifelse(matches$bResult > 0, c("blue"), c("red"))

# matches$YearSeason <- as.character(matches$Year)
matches$YearSeason <- ifelse(matches$Season == "Spring", paste("01/01/", as.character(matches$Year), sep=""), paste("01/06/", as.character(matches$Year), sep=""))

str(matches$YearSeason)
summary(matches$Season)


matches$YearSeason <- as.Date(matches$YearSeason, format = "%d/%m/%Y")

summary(matches$YearSeason)
```

```{r echo=FALSE, warning=FALSE, results = "hide", "0.6 Making subsets"}
# It gets messy very easily so extra dataset instead of extra variable:
matches_seperated <- matches %>%
  separate(golddiff, into = c(create_column_names(gamelength, "gold diff in minute")), sep = "\\,")

# may be interesting to make a red vs blue won dataset.

blueset <- subset(matches, winner == 'blue')
redset <- subset(matches, winner == 'red')
```

# 1. EXPLORING THE DATA SET (univariate and some bivariate).

## 1.1 First exploratory plots (see below graph for analysis)

```{r}
# Get the names
names(matches)
# 58 original vars

str(matches)
# So there are 185 teams --> seems like a lot? Do teams change from name? Do members change?

# this runs a whole summary (lot of work):
summary(matches)

table(matches$winner)
# ---> so blue wins (a lot) more
```

Observations:
Regional competition is very small: 143... exclude those if they behave very differently

There are two seasons, so we seem to have Year*Season = 4*2=8 time values

Game length mean and median are very close: so it is not scewed.

Most notable is the choice for Maokai for top red--> he is a lot more favorite for red than blue...

```{r echo=FALSE, message=FALSE, warning=FALSE, results = 'hide'}
# make a table of RG Case --> it is exactly the same, so that seems correct
table(matches$winner)
#it is exactly the same, so that seems correct
```
## 1.2 First exploratory plots: length of the game and the different leagues, types & timeframes(seasons)

How are the leagues distributed (univariate)?  
```{r echo=FALSE, message=FALSE, warning=FALSE, "1.2 First exploratory plots"}
qplot(x = League, data = matches,
      xlab = 'league',
      ylab = 'Number of games from sample',
      color = I('black'), fill = I('blue'))
```
A lot of variation, with clear distinction. The game should function the same across the leagues.

How are the types distributed (univariate)?  
```{r echo=FALSE, message=FALSE, warning=FALSE, "1.2 First exploratory plots"}
qplot(x = Type, data = matches,
      xlab = 'league',
      ylab = 'Number of games from sample',
      color = I('black'), fill = I('yellow'))
summary(matches$Type)
```
Huge difference between season and regional --> see if regional behaves differently, maybe exclude from sample

How are the seasons distributed (univariate)?  
```{r echo=FALSE, message=FALSE, warning=FALSE, "1.2 First exploratory plots"}
qplot(x = Type, data = matches,
      xlab = 'league',
      ylab = 'Number of games from sample',
      color = I('black'), fill = I('yellow'))
summary(matches$Type)
```





How is the time distributed (univariate)?  
```{r echo=FALSE, message=FALSE, warning=FALSE, "1.2 First exploratory plots"}
qplot(x = gamelength, data = matches,
      xlab = 'length of game',
      ylab = 'Number of games from sample',
      color = I('black'), fill = I('purple')) +
  scale_x_continuous()
```
Like we saw from the summary, the data is nicely distributed. It is also clear that it is almost impossible to win this game in the first 20 minutes in the professional spfere.

How are the matches distributed of the seasons (univariate)?  
```{r echo=FALSE, message=FALSE, warning=FALSE, "1.2 First exploratory plots"}
qplot(x = YearSeason, data = matches,
      xlab = 'Season',
      ylab = 'Number of games from sample',
      color = I('black'), fill = I('purple'))
```
Wow, there is way more data from later timeslots...

How is the time distributed over the different leagues (bivar)?

```{r}

```

