---
title: "Process the league of legends dataset"
output: html_notebook
---

## Global configuration variables to control the data in the generated file
```{r}
 shouldDiscardTimeSeriesColumnsFromProcessedFile <- FALSE
 processedFileNameWithExtension <- "processed_matches.csv"
```

## Process the current data and get:

1. Number of game elements (towers, inhibitors, dragons and barons) killed or destroyed by red and blue teams in the columns bTowersNum	bInhibsNum	bDragonsNum	bBaronsNum	bHeraldsNum	rTowersNum	rInhibsNum	rDragonsNum	rBaronsNum	rHeraldsNum

2. The time when the game element (towers, inhibitors, dragons and barons) was destroyed by the red and the blue teams. This value is 1000 if there was no element destroyed by that team. The columns created are bTowersFirstDestroyedTime,	rTowersFirstDestroyedTime,		bInhibsFirstDestroyedTime,	rInhibsFirstDestroyedTime,		bDragonsFirstDestroyedTime,	rDragonsFirstDestroyedTime

3. The team that first killed pr destroyed a game element (towers, inhibitors, dragons and barons). This value is 1 if it was performed by the Blue team and 0 if performed by the Red team. These columns are TowersFirstKillBy,	InhibsFirstKillBy, DragonsFirstKillBy
```{r}
library(stringr)

setwd("H:\\KE 5107 Projects\\code\\KE5107_LeagueofLegends")
matchesData = read.csv("data/matches.csv")
gameElements <- c("Towers", "Inhibs", "Dragons", "Barons") 

# Prepends r to the item
prependRed <- function(p) {
  paste("r", p, sep="")
}

# Prepends b to the item
prependBlue <- function(p) {
  paste("b", p, sep="")
}

# Get the smallest number from a series of numbers
getSmallestNumber <- function(value) {
  if (value != "[]") {
    # Get all the number strings from the list
    allNumbers <- str_extract_all(value, "[0-9]*[.][0-9]*")
    
    # Get the number strings in a list
    allNumbersList <- unlist(allNumbers)
    
    # Convert each item to a number
    allNumbersList <- as.numeric(allNumbersList)
    
    # Sort all the numbers
    allNumbersListSorted <- allNumbersList[order(allNumbersList)]
    
    # Get the first number
    return(allNumbersListSorted[1])
  }
  
  # A high value if the no number is not present
  return(10000)
}

# Get the Kills columns for blue and red teams
blueKillsColumns <- lapply(gameElements, prependBlue)
redKillsColumns <- lapply(gameElements, prependRed)

# Combine the Kills columns
allKillsColumns <- c(blueKillsColumns, redKillsColumns)

for(i in 1:length(allKillsColumns)) {
  columnName <- unlist(allKillsColumns[i])
  outputColumnName <- paste(columnName, "Num", sep="")
  
  # Each item is enclosed in its own square brackets. Note that there is an extra [ to contain all the elements which needs to be removed
  matchesData[,outputColumnName] <- str_count(matchesData[,columnName], "\\[")-1
}

firstKillElements <- c("Towers", "Inhibs", "Dragons", "Barons") 
for(i in 1: length(firstKillElements)){
  columnName <- unlist(firstKillElements[i])
  
  # Get the red and blue specific column names
  redColumnName <- paste("r", columnName,sep="")
  blueColumnName <- paste("b", columnName,sep="")
  
  # Get the output column name to store the first destroyed time
  redOutputColumnName <- paste("r", columnName, "FirstDestroyedTime",sep="")
  blueOutputColumnName <- paste("b", columnName, "FirstDestroyedTime",sep="")
  
  # Get the first number from the list of numbers
  
  matchesData[,blueOutputColumnName] <- unlist(lapply(as.character(matchesData[,blueColumnName]), getSmallestNumber))
  
  matchesData[,redOutputColumnName] <- unlist(lapply(as.character(matchesData[,redColumnName]), getSmallestNumber))
  
  # convert the first kill times to numeric for comparison
  matchesData[,blueOutputColumnName] <- as.numeric(matchesData[,blueOutputColumnName])
  matchesData[,redOutputColumnName] <- as.numeric(matchesData[,redOutputColumnName])
  
  # Get the name of the output column
  firstKillByColumnName <- paste(columnName, "FirstKillBy",sep="")
  
  matchesData[,firstKillByColumnName] <- as.numeric(matchesData[,blueOutputColumnName] < matchesData[,redOutputColumnName])
}

```
## Code to remove existing columns
```{r}
# remove the gold diff column as it has redundant information
matchesData$golddiff <- NULL

# remove the rResults column as it too has redundant information
matchesData$rResult <- NULL

if(shouldDiscardTimeSeriesColumnsFromProcessedFile){
  timeSeriesColumnNames <- c("goldblue", "bKills", "bTowers", "bInhibs", "bDragons", "bBarons", "bHeralds", "goldred", "rKills", "rTowers", "rInhibs", "rDragons", "rBarons", "rHeralds", "goldblueTop", "goldblueJungle", "goldblueMiddle", "goldblueADC", "goldblueSupport", "goldredTop", "goldredJungle", "goldredMiddle", "goldredADC", "goldredSupport")
  
  for(i in 1:length(timeSeriesColumnNames)){
    columnName <- unlist(timeSeriesColumnNames[i])
    matchesData[,columnName] <- NULL
  }
}

```
## Insert code to do other processing of the matches data
```{r}
```
## Write the current contents of the matchesData dataframe into a CSV in the current folder
```{r}
write.csv(matchesData, processedFileNameWithExtension, row.names = FALSE)
```
